---
title: "Basics"
date: "`r lubridate::now()`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>"
)
```

## Premble

The code presented here is inspired by the supplement to [“Identifying fishing grounds from vessel tracks: model-based inference for small scale fisheries” by Tania Mendo, Sophie Smout, Theoni Photopoulou and Mark James (2019) Royal Society Open Science](https://doi.org/10.1098/rsos.191161). The supplementary document provides a code for "five different methods for identifying hauling versus non-hauling activities in small scale fishing vessels from their movement trajectories."

A [sample dataset](https://doi.org/10.5061/dryad.k80bp46) of movement data collected every 60sec from 5 trips by 5 different small scale fishing vessels using creels is provided and used here (has been included in the {ramb}-package).

The objective here is to provide a more structured code than provided in the supplement, relying fully on tidyverse lingo including usage of the {purrr} map-function family, rather than loops. Those details are though hidden within the function-calls, the aim of the default output those that are needed for vessel track behavioural postprocessing analysis.

```{r}
library(ramb)
library(tidyverse)
library(lubridate)
```

```{r}
creel.change <- 
  creel %>% 
  group_by(id) %>% 
  mutate(change = rb_event(behaviour)) %>% 
  group_by(id, change, behaviour) %>% 
  summarise(t1 = min(time),
            t2 = max(time),
            .groups = "drop") %>% 
  # make a continuum, not loose the last point
  group_by(id) %>% 
  mutate(t2 = lead(t1)) %>% 
  ungroup()
p.base <- 
  creel.change %>% 
  ggplot() +
  theme_bw() +
  geom_rect(aes(xmin = t1, xmax = t2, ymin = -Inf, ymax = Inf,
                fill = behaviour),
            show.legend = FALSE) +
  facet_wrap(~ id, scales = "free") +
  scale_fill_manual(values = c("steaming" = "grey",
                               "hauling" = "pink",
                               "shooting" = "green")) +
  ggnewscale::new_scale_fill()
```

## A restructured code

### Model 1: Gaussian mixture model

... not shown.

### Model 2: Trip-based Gaussian mixture model

```{r}
d <-
  creel %>%
  group_by(id) %>%
  ramb::rb_gaussian() %>% 
  mutate(Fishing = ifelse(speed <= threshold.upper, TRUE, FALSE))

d %>%
  count(behaviour, Fishing) %>%
  spread(Fishing, n) %>%
  knitr::kable()
d %>%
  ggplot(aes(speed, fill = Fishing)) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(~ id, scales = "free_y") +
  theme(legend.position = c(0.8, 0.2)) +
  scale_fill_brewer(palette = "Set1") +
  labs(x = "Speed [kt]", y = "Pings")
p.base +
  geom_point(data = d,
             aes(time, speed, colour = Fishing),
             size = 0.5) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Time", y = "Speed [kt]") +
  theme(legend.position = c(0.8, 0.2))
```

### Model 3: Trip-based Binary Clustering using Gaussian mixture models on a trip-by-trip basis

```{r}
d <- 
  creel %>% 
  group_by(id) %>% 
  ramb::rb_gaussian_binary_clustering() %>% 
  mutate(Fishing = ifelse(A %in% 1:2, TRUE, FALSE))
d %>%
  count(behaviour, A) %>%
  spread(A, n) %>%
  knitr::kable()
d %>%
  count(behaviour, Fishing) %>%
  spread(Fishing, n) %>%
  knitr::kable()
d %>%
  ggplot(aes(speed, turn, colour = factor(A))) +
  theme_bw() +
  geom_point(size = 1) +
  facet_wrap(~ id) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Speed [nm]", y = "Turn", colour = "clusters") +
  theme(legend.position = c(0.8, 0.25))
d %>%
  ggplot(aes(time, speed)) +
  theme_bw() +
  geom_point(aes(colour = factor(A)),
             size = 1) +
  facet_wrap(~ id, scales = "free") +
  theme(legend.position = c(0.8, 0.2)) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Time", y = "Speed [kt]")
p.base +
  geom_point(data = d,
             aes(time, speed, colour = Fishing),
             size = 0.5) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Time", y = "Speed [kt]") +
  theme(legend.position = c(0.8, 0.2))
d %>%
  ggplot(aes(time, turn)) +
  theme_bw() +
  geom_point(aes(colour = factor(A)),
             size = 1) +
  facet_wrap(~ id, scales = "free") +
  theme(legend.position = c(0.8, 0.2)) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Time", y = "Turn")
p.base +
  geom_point(data = d,
             aes(time, turn, colour = Fishing),
             size = 0.5) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Time", y = "Turn") +
  theme(legend.position = c(0.8, 0.2))
d %>%
  arrange(desc(A)) %>%
  ggplot(aes(lon, lat, colour = factor(A))) +
  theme_bw() +
  geom_point(size = 1) +
  facet_wrap(~ id, scales = "free") +
  coord_quickmap() +
  labs(x = NULL, y = NULL, colour = "clusters") +
  scale_colour_brewer(palette = "Set1") +
  theme(legend.position = c(0.8, 0.25))  +
  scale_x_continuous(NULL, NULL) +
  scale_y_continuous(NULL, NULL)
d %>%
  arrange(desc(A)) %>%
  ggplot(aes(lon, lat, colour = Fishing)) +
  theme_bw() +
  geom_point(size = 1) +
  facet_wrap(~ id, scales = "free") +
  coord_quickmap() +
  labs(colour = "Fishing") +
  scale_colour_brewer(palette = "Set1") +
  theme(legend.position = c(0.8, 0.25)) +
  scale_x_continuous(NULL, NULL) +
  scale_y_continuous(NULL, NULL)
```

### Model 4: Hidden Markov model with speed only

where is the speed here?


```{r}
d <-
  creel %>%
  ramb::rb_hidden_markov_step()
d %>%
  count(behaviour, vit) %>%
  spread(vit, n) %>%
  knitr::kable()
p.base +
  geom_point(data = d,
             aes(time, step, colour = factor(vit)),
             size = 0.5) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Time", y = "Step [m]") +
  theme(legend.position = c(0.8, 0.2))
d %>%
  mutate(id = paste(id, behaviour)) %>%
  ggplot(aes(time, step, colour = factor(vit))) +
  geom_point() +
  facet_wrap(~id, scales = "free", ncol = 3) +
  scale_colour_brewer(palette = "Set1")
# why not on speed??
```

### Model 5: Hidden Markov Model with speed and turning angle

```{r}
d <- 
  creel %>% 
  ramb::rb_hidden_markov_step_and_turn()
p.base +
  geom_point(data = d,
             aes(time, step, colour = factor(vit)),
             size = 0.5) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Time", y = "Step [m]") +
  theme(legend.position = c(0.8, 0.2))
d %>% 
  mutate(id = paste(id, behaviour)) %>%
  ggplot(aes(time, step, colour = factor(vit))) +
  geom_point() +
  facet_wrap(~id, scales = "free", ncol = 3) +
  scale_colour_brewer(palette = "Set1")
```


