---
title: "Basics"
date: "`r lubridate::now()`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>"
)
```

```{r}
library(sf)
library(mapview)
library(mapdeck)
library(lubridate)
library(ramb)
library(tidyverse) # alway load this last (because of potential function name conflicts)
```

```{r, echo = FALSE}
fil <- "~/r/Pakkar2/ramb/TOPSECRET.R"
if(file.exists(fil)) source(fil) 
```

## Vessel in harbour (A point in polygon issue)

Example data: Icelandic survey tracks and some harbour polygons

```{r}
harbours.sf <- 
  read_sf("ftp://ftp.hafro.is/pub/data/shapes/harbours.gpkg") %>% 
         # only harbours in Iceland
  filter(iso2a == "IS",
         # filter out some smaller harbours, cause a nuisance downstream
         !hid %in% c("HRI", "ASS", "HAU", "GRE", "MJH", "MJO", "AED", "HJA")) %>% 
  select(hid)

tracks.sf <- 
  read_is_survey_tracks() %>% 
  mutate(speed = ifelse(speed > 12, 12, speed)) %>% 
  st_as_sf(coords = c("lon", "lat"),
           crs = 4326,
           remove = FALSE) %>% 
  st_join(harbours.sf) %>% 
  mutate(inharbour = ifelse(!is.na(hid), TRUE, FALSE))
```

```{r}
bb <- st_bbox(c(xmin = -22.75, ymin = 64.05, 
                xmax = -21.65, ymax = 64.20), 
              crs = 4326)

mapdeck(location = c(-22.0, 64.1), zoom = 9) %>% 
  add_polygon(data = harbours.sf, fill_colour = "#FF000095", tooltip = "hid", update_view = FALSE) %>% 
  add_scatterplot(data = tracks.sf %>% st_crop(bb) %>% arrange(inharbour), 
                  fill_colour  = "inharbour", radius = 100, update_view = FALSE)
```

## Trip identification

One could define a trip as all the consecutive points while not in harbour. Determining if points are inside or outside a harbour have been dealt with above. The {data.table} \code{rleid} function deals with generating a run-length type group id

```{r}
# rleid(cumsum(!mydata$Info == "3D"))
# try a conditional rleid(cumsum(!inharbour == FALSE))
tracks <- 
  tracks.sf %>% 
  st_drop_geometry() %>% 
  # here we do not want to count harbour stay as a separate trip
  group_by(vid) %>% 
  mutate(.gr0 = data.table::rleid(inharbour)) %>% 
  group_by(vid, inharbour) %>% 
  mutate(tripid = data.table::rleid(.gr0)) %>% 
  ungroup() %>% 
  select(-.gr0)
tracks %>% 
  filter(vid == 1579) %>% 
  ggplot() +
  theme_void() +
  geom_path(aes(lon, lat, colour = as_factor(tripid)), lwd = 1) +
  geom_path(data = geo::island, aes(lon, lat), colour = "grey", lwd = 1) +
  scale_colour_brewer(palette = "Set1") +
  coord_quickmap() +
  labs(colour = "Trip no") +
  theme(legend.position = "bottom")
```

We can take a peek at the trip info's:

```{r}
tracks %>% 
  filter(vid == 1579) %>% 
  group_by(tripid) %>% 
  mutate(distance = geo::arcdist(lat, lon, lead(lat), lead(lon)),
         distance = replace_na(distance, 0)) %>% 
  summarise(n = n(),
            distance = sum(distance), # nautical miles
            t1 = min(time),
            t2 = max(time),
            das = as.numeric(difftime(t2, t1, units = "days")))
```

We see here that trip 5 is very short, the problem lies with having a wacky point, that needs to be fixed upstream:
```{r}
tracks %>% 
  filter(vid == 1579) %>% 
  slice(2370:2380) %>% 
  mutate(dist = geo::arcdist(lat, lon, lead(lat), lead(lon))) %>% 
  knitr::kable(digits = 2)
```


```{r, eval = FALSE, echo = FALSE}
d <- 
  tracks %>% 
  filter(vid == 1579) %>% 
  st_as_sf(coords = c("lon", "lat"),
           crs = 4326)
d2 <- 
  d %>% 
  st_drop_geometry() %>% 
  select(tripid) %>% 
  distinct()
pal <- c("#A6CEE3FF", "#1F78B4FF", "#B2DF8AFF", "#33A02CFF", "#FB9A99FF", 
        "#E31A1CFF", "#FDBF6FFF", "#FF7F00FF", "#CAB2D6FF", "#6A3D9AFF", 
        "#FFFF99FF", "#B15928FF")
d2$pal <- c(pal, pal)[1:nrow(d2)]
d <- 
  d %>% 
  left_join(d2)
mapdeck(style = mapdeck_style("dark")) %>% 
  add_scatterplot(data = d, 
                  fill_colour = "pal", 
                  tooltip = "tripid", 
                  legend = FALSE,
                  radius = 250,
                  radius_min_pixels = 2,
                  radius_max_pixels = 10)
```



## Calculation of trip length and duration, including cumulatives

```{r}
tracks <- 
  tracks %>% 
  group_by(vid, tripid) %>% 
  mutate(dist = geo::arcdist(lat, lon, lead(lat), lead(lon)),
         dist = replace_na(dist, 0),
         cdist = cumsum(dist),
         dura = as.numeric(difftime(lead(time), time, units = "hours")),
         dura = replace_na(dura, 0),
         cdura = cumsum(dura)) %>% 
  ungroup()
tracks %>% 
  filter(vid == 1579,
         !inharbour) %>% 
  ggplot() +
  theme_void() +
  geom_point(aes(lon, lat, colour = cdist)) +
  geom_path(aes(lon, lat, group = tripid), colour = "grey") +
  geom_polygon(data = geo::island, aes(lon, lat), fill = "grey") +
  scale_colour_viridis_c(option = "B", direction = -1) +
  coord_quickmap()
```


```{r, eval = FALSE}
# checkout: geosphere::distHaversine
tracks %>% 
  mutate(dist2 = geosphere::distHaversine(c(lon, lat)))
```


## remains

```{r}
track <-
  read_csv("https://raw.githubusercontent.com/ices-eg/WKSSFGEO/main/example_data_AIS.csv") %>% 
  # reduce typing in downstream code
  rename(vid = vessel_id, time = time_stamp) %>% 
  # unique rows
  distinct() %>% 
  # remove duplicate time
  distinct(vid, time, .keep_all = TRUE) %>% 
  # # remove same positions (this is not recommended could theoretically be possible
  # distinct(vid, lon, lat, .keep_all = TRUE)
  arrange(vid, time) %>% 
  group_by(vid) %>% 
  mutate(duration = difftime(time, lag(time), units = "hours"),
         # warnings here are just because of the first datapoint
         distance = geo::arcdist(lat, lon, lag(lat), lag(lon), scale = "nmi"),
         speed_derived = distance / as.numeric(duration))
```


```{r}
track %>% 
  sample_n(1e4) %>% 
  ggplot(aes(speed, speed_derived)) +
  geom_point(alpha = 0.01, size = 1) +
  geom_abline(colour = "red") +
  coord_equal()
```

