---
title: "Predicting gear width"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE,
  collapse = TRUE,
  error = TRUE,
  comment = "#>"
)
```

## Preamble

The ICES VMS datacall request metrics on gear width of towed gear. Routinely the gear width is obtained from 'the Benthis' model. In this model gear width is a function of gear type and vessel length or power (kW), the model being either linear or a power function. Here a code snipped is provided that may be of use for predicting gear width.

## Data preparation

Here will use the eflalo dataset from {vmstools}

```{r}
library(ramb)
library(tidyverse)
library(vmstools)  # to get access to example data
data("eflalo")
lb <- 
  eflalo |> 
  as_tibble() |> 
  mutate(metier5 = rb_met5_from6(LE_MET_level6)) |> 
  select(length = VE_LEN, kw = VE_KW, metier5, metier6 = LE_MET_level6) |> 
  distinct() |> 
  mutate(.rid = 1:n(),
         .before = length)
```

## Coding within a flow

A simple pipe flow:

```{r}
lb1 <- 
  lb |> 
  left_join(metier5_benthis_lookup,
          by = join_by(metier5)) |> 
  left_join(benthis_parameters |> select(benthis_metier:variable),
            by = join_by(benthis_metier)) |> 
  mutate(width = 
           case_when(
             model == "power" & variable == "avg_kw" ~    a * kw^b,
             model == "power" & variable == "avg_oal" ~   a * length^b,
             model == "linear" & variable == "avg_oal" ~  a * length + b,
             .default = NA))
```

## Using a function

Using the function in {ramb}

```{r}

lb2 <- 
  lb |> 
  mutate(width = rb_benthis_width(metier5, length, kw))
identical(lb1$width, lb2$width)
```

## Comparision with ICES VMS datacall flow

Just a double check:

```{r}
source("~/stasi/ices/ICES-VMS-and-Logbook-Data-Call/0_global_functions.R")
lb3 <- lb
lb3$width <- add_gearwidth(lb3, "metier6", "length", "kw")
identical(lb2$width / 1000, lb3$width)
near(lb2$width / 1000, lb3$width) |> table(useNA = "ifany")
```

