---
title: "Interpolations"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE,
  collapse = TRUE,
  error = TRUE,
  comment = "#>"
)
```

**NOTE**: 

* just a templated document so far, need to code
* check out and compare jepol's function and the traipse::track_intermediate

```{r, eval = FALSE}
library(data.table)
library(sf)
library(traipse)
library(tidyverse)
# code snippet example, needs further work
tmp <-
  out.eh %>% 
  filter(vessel_id == 1) %>% 
  group_by(vessel_id, tid) %>% 
  mutate(x = track_intermediate(lon, lat,
                                date = time_stamp, duration = 2)) %>% 
  ungroup()
tmp %>% 
  unnest(x)

hb <- 
  ramb::read_is_harbours() %>%
  filter(iso2a == "IS",
         # filter out some smaller harbours, cause a nuisance downstream
         !hid %in% c("HRI", "ASS", "HAU", "GRE", "MJH", "MJO", "AED", "HJA")) %>% 
  mutate(hid = 1:n())
d <- 
  ramb::read_is_survey_tracks() %>% 
  mutate(.rid = 1:n()) %>% 
  arrange(vid, time) %>% 
  select(.rid, everything()) %>% 
  st_as_sf(coords = c("lon", "lat"),
           crs = 4326,
           remove = FALSE) %>% 
  st_join(hb %>% select(hid)) %>% 
  st_drop_geometry() %>% 
  # here check for wacky points
  group_by(vid) %>% 
  mutate(pings = n(),
         sd = ifelse(pings > 3, track_distance(lon, lat), NA_real_),
         ss = ifelse(pings > 3, track_speed(lon, lat, time), NA_real_),
         st = track_time(time),
         duration = sum(st) / 60,      # duration in minutues
         whacky = ramb::rb_whacky_speed(lon, lat, time)) %>% 
  ungroup()
# define trips, then interpolate
d <- 
  d %>% 
  group_by(vid) %>% 
  mutate(inh = !is.na(hid),
         tid = ramb::rb_trip(inh)) %>% 
  ungroup()
d %>% 
  filter(vid == 1131, tid == 1) %>% 
  filter(between(lon, -22, -10),
         between(lat, 66.25, 70)) %>% 
  group_by(vid, tid) %>% 
  mutate(x = track_intermediate(lon, lat,
                                date = time, duration = 10*60)) %>% 
  ungroup() %>% 
  unnest(x) %>% 
  ggplot() +
  geom_point(aes(int_x, int_y), colour = "blue", size = 0.7) +
  geom_point(aes(lon, lat), colour = "red", size = 0.3) +
  coord_quickmap()
  ```


## Comparison on trip calculation

```{r}
library(data.table)
library(traipse)
library(tidyverse)
```

Data:

```{r}
hbs  <- readRDS("~/ass/2021/WKSSFGEO/github/WKSSFGEO/data/harbours.rds")
githubURL1 <- "https://raw.githubusercontent.com/ices-eg/WKSSFGEO/main/data-examples/example_data_AIS.csv"
d <- 
  readr::read_csv(githubURL1) %>% 
  unique(by = c("vessel_id", "time_stamp")) %>% # Remove duplicates
  mutate(vessel_id = str_sub(vessel_id, 4) %>% as.integer(),
         .rid = 1:n())
```


Jeppe:

```{r}
devtools::source_url("https://raw.githubusercontent.com/ices-eg/WKSSFGEO/main/R-dev/jepol/define_trips_pol.R")
devtools::source_url("https://raw.githubusercontent.com/ices-eg/WKSSFGEO/main/R-dev/jepol/add_harbours.R")
# devtools::source_url("https://raw.githubusercontent.com/ices-eg/WKSSFGEO/main/R-dev/jepol/interpolate_ais.R")
system.time(out.jp <-
              readr::read_csv(githubURL1) %>% 
              unique(by = c("vessel_id", "time_stamp")) %>% # Remove duplicates
              mutate(vessel_id = str_sub(vessel_id, 4) %>% as.integer(),
                     .rid = 1:n()) %>% 
              setDT() %>% 
              add_harbours(hbs) %>% 
              define_trips_pol(min_dur = 0.0001, max_dur = 50e6, 
                               split_trips = FALSE, preserve_all = TRUE))
```

Einar:

```{r}
rb_trip <- function(x) {
  tibble::tibble(x = x) %>% 
    dplyr::mutate(tid = dplyr::if_else(x != dplyr::lag(x), 1L, 0L, 1L),
                  tid = ifelse(x, -tid, tid)) %>% 
    dplyr::group_by(x) %>% 
    dplyr::mutate(tid = cumsum(tid)) %>% 
    dplyr::ungroup(x) %>% 
    dplyr::pull(tid)
}

system.time(out.eh <- 
              readr::read_csv(githubURL1) %>% 
              unique(by = c("vessel_id", "time_stamp")) %>% # Remove duplicates
              mutate(vessel_id = str_sub(vessel_id, 4) %>% as.integer(),
                     .rid = 1:n()) %>% 
              sf::st_as_sf(coords = c("lon", "lat"),
                           crs = 4326,
                           remove = FALSE) %>% 
              sf::st_join(hbs %>% mutate(hid = 1:n()) %>% select(hid)) %>% 
              sf::st_drop_geometry() %>% 
              group_by(vessel_id) %>% 
              mutate(inh = !is.na(hid),
                     tid = rb_trip(inh)) %>% 
              ungroup())


# Pad harbour id, adding harbour id left to the first ping of a trip and harbour id entered
#  to the last ping of the trip.
rb_pad_harbour <- function(tid, hid) {
  tibble::tibble(.tid = tid,
                 .hid = hid) %>% 
    dplyr::mutate(.hid = dplyr::case_when(.tid !=  dplyr::lag(.tid) & .tid > 0 ~  dplyr::lag(.hid),
                                          .tid != dplyr::lead(.tid) & .tid > 0 ~ dplyr::lead(.hid),
                                          TRUE ~ .hid)) %>% 
    dplyr::pull(.hid)
}
out.eh %>% 
  group_by(vessel_id) %>% 
  mutate(hid = rb_pad_harbour(tid, hid)) %>% 
  filter(tid > 0) %>% 
  group_by(vessel_id, tid) %>% 
  slice_tail(n = 1) %>% 
  view()
```

Number of records:

```{r}
tibble(original = nrow(d),
       jp = nrow(out.jp),
       eh = nrow(out.eh))
```

Any vessel difference?:

```{r}
full_join(out.eh %>% count(vessel_id, name = "n.eh"),
          out.jp %>% count(vessel_id, name = "n.jp"))
```

So vessels 5 and 8 not returned in the jp-workflow.

```{r}
out.eh %>% 
  filter(vessel_id %in% c(5, 8)) %>% 
  count(vessel_id, inh)
```

Reason seemms the jp-flow does not return records if all pings are in harbour. This may be intentional - should possibly be mentioned in the help file that this takes place.

Number of pings per trip:

```{r}
full_join(out.eh %>% mutate(tid = ifelse(tid < 0, NA, tid)) %>% count(vessel_id, tid, name = "n.eh"),
          out.jp %>% separate(trip_id, c("dummy", "tid"), convert = TRUE) %>% count(vessel_id, tid, name = "n.jp")) %>% 
  mutate(diff = n.eh - n.jp) %>% 
  knitr::kable()
```

So difference is in all cases except vessel 7, trip 8 one less ping in the eh-algorithm compared with the jp-algoritm. 

```{r}
tmp <-
  out.eh %>%
  select(.rid, vessel_id, time_stamp, hid.eh = hid, inh, tid.eh = tid) %>%
  full_join(out.jp %>% as_tibble() %>% select(.rid, hid.jp = SI_HARB, tid.jp = trip_id) %>% mutate(in.jp = TRUE)) %>%
  mutate(in.jp = replace_na(in.jp, TRUE)) %>%
  separate(tid.jp, c("dummy", "tid.jp"), convert = TRUE) %>%
  select(-dummy)
```

The reason for the difference of one is that the jp-algorithm includes the first harbour ping at the end of a trip as being part of the trip. Case in point:

```{r}
tmp %>% 
  filter(vessel_id == 1, tid.jp == 1) %>% 
  tail()
```

What about vessel 7, trip 8? 

```{r}
tmp %>% 
  filter(vessel_id == 7, tid.jp == 8) %>% 
  data.table()
```

Seems like the last record for this trip did not terminate in a harbour, as illustrated here (actually it does not even start in a harbour):

```{r}
out.eh %>% 
  filter(vessel_id == 7, .rid > 137770) %>% 
  sf::st_as_sf(coords = c("lon", "lat"),
               crs = 4326) %>% 
  mutate(tid = as.character(tid)) %>% 
  mapview::mapview(zcol = "tid")
```
