---
title: "MFRI workflow"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE,
  collapse = TRUE,
  error = TRUE,
  comment = "#>"
)
```

## Preamble

```{r}
library(data.table)
library(ramb)
library(sf)
library(mapdeck)
library(traipse)
library(tidyverse)
source("~/r/Pakkar2/ramb/TOPSECRET.R")
```

## Concept

We have one day of track (1 minute interval) for three vesssels (lon and lat not inluded) and three logbook records (two from one vessel, one from another vessel - one vessel has no logbook record). The logbooks recordings are among other things: time of setting (to), time of start hauling (t1) and time of end hauling (t2).

Want to assign ...

```{r}
stk <- 
  expand_grid(vid = c(1000, 1001, 1002),
              time = seq(ymd_hms("2022-09-22 00:00:00"), 
                         ymd_hms("2022-09-22 23:59:59"), 
                         by = "min"))
lgs <- 
  tibble(vid = c(1000, 1000, 1001),
         visir = c(-1, -2, -3),
         start_setting = c(ymd_hms("2022-09-22 10:00:00"), ymd_hms("2022-09-22 15:00:00"), ymd_hms("2022-09-22 10:00:00")),
         start_haul = c(ymd_hms("2022-09-22 11:00:00"), ymd_hms("2022-09-22 16:00:00"), ymd_hms("2022-09-22 11:30:00")),
         end_haul = c(ymd_hms("2022-09-22 14:00:00"), ymd_hms("2022-09-22 18:30:00"), ymd_hms("2022-09-22 18:00:00")))
lgs
stk %>% 
  rb_interval_id2(lgs, vid, time, start_haul, end_haul, visir) %>% 
  ggplot(aes(time, factor(vid), fill = factor(.id))) +
  geom_tile()
```

## Longline

```{r}
library(omar)
source("R/rb_workflow.R")
con <- connect_mar()
```

```{r}
YEARS <- 2021
vessels <- 
  omar::vid_registry(con, standardize = TRUE) %>% 
  # no dummy vessels
  filter(!vid %in% c(0, 1, 3:5),
         !is.na(vessel))
gears <-
  tbl_mar(con, "ops$einarhj.gid_orri_plus") %>% 
  select(gid = veidarfaeri, gclass = gid2) %>% 
  collect(n = Inf) %>% 
  mutate(#description = ifelse(gid == 92, "G.halibut net", description),
    gid = as.integer(gid),
    gclass = as.integer(gclass))
```


### Logbooks

PROCESSING STEPS:

1. Get and merge logbook and landings data
* gear correction part of the process
2. Lump some gears
3. Cap effort and end of action
4. Mesh size "corrections"
5. Set gear width proxy
6. Set gear class


8. Match vid with mobileid in stk - pending
9. Add vessel information - only needed for ICES datacall - pending
10. Add metier - only needed for ICES datacall - pending
11. ICES rectangles - only needed for ICES datacall - pending
12. Anonymize vid - only needed for ICES datacall - pending


```{r}
lb <- 
  vessels %>% 
  # need to carefully review the changes, went from mar to omar in the function call
  #  the old process is in the rb_logbooks_2022
  rb_logbook(YEARS)
lb %>% count(gid)

 # only vessels that are longlining, and that always fill out t0, t1, t2
lb %>% 
  mutate(crit = ifelse(!is.na(t0) & !is.na(t1) & !is.na(t2), TRUE, FALSE)) %>% 
  group_by(vid) %>% 
  summarise(n.settings = n(),
            gid1 = sum(gid == 1),
            crit = sum(crit),
            .groups = "drop") %>% 
  filter(n.settings == gid1) %>% 
  filter(n.settings == crit) %>% 
  arrange(n.settings) %>% 
  filter(n.settings > 10) %>% 
  pull(vid) ->
  VIDs
lb <- 
  lb %>% 
  filter(vid %in% VIDs) %>% 
  # get the mobileid
  #  Here we get into the problem that we have multiple mid for a vid
  #  ergo: repeated records in the logbooks
  left_join(tbl_mar(con, "ops$einarhj.MID_VID_ICELANDIC_20220418") %>% 
              select(mid, vid) %>% 
              collect(n = Inf),
            by = "vid") %>% 
  select(vid, mid, visir, gid, date, t0, t1, t2, effort, catch, fj_kroka)
```

```{r}
MIDs <- 
  lb %>% 
  pull(mid) %>% 
  unique()
stk.raw <- 
  omar::stk_trail(con, YEARS) %>% 
  filter(mid %in% MIDs) %>% 
  collect(n = Inf) %>% 
  left_join(lb %>% select(vid, mid) %>% distinct()) %>% 
  mutate(time = round_date(time, unit = "seconds"))
# without extrapolations
stk.tmp <- 
  stk.raw %>% 
  rb_interval_id2(lb, vid, time, t1, t2, visir) %>% 
  rename(visir = .id) %>% 
  mutate(fishing = if_else(!is.na(visir),TRUE, FALSE, FALSE)) 
stk.tmp %>% 
  st_as_sf(coords = c("lon", "lat"),
           crs = 4326) %>% 
  mapdeck %>% 
  add_scatterplot(fill_colour = "fishing",
                  radius = 500,
                  palette = "inferno")
```


```{r}
tmp <-
  stk.tmp %>% 
  filter(fishing) %>%
  # try also filtering speed
  filter(between(speed, 0.375, 2.750)) %>% 
  select(vid, visir, lon, lat) %>% 
  st_as_sf(coords = c("lon", "lat"),
           crs = 4326) %>% 
  group_by(vid, visir) %>% 
  summarise(do_union = FALSE,
            .groups = "drop") %>% 
  st_cast("LINESTRING")
tmp %>% 
  slice(1:10) %>% 
  select(visir) %>% 
  mutate(visir = as_factor(visir)) %>% 
  plot()

tmp %>% 
  #sample_n(1000) %>% 
  mutate(length = as.numeric(st_length(.))) %>% 
  st_drop_geometry() %>% 
  left_join(lb %>% select(vid, visir, effort)) %>% 
  filter(between(effort, 13000, 19400) | between(effort, 36000, 46000)) %>% 
  mutate(small = ifelse(effort < 20000, TRUE, FALSE)) %>% 
  mutate(n = effort / length) %>% 
  mutate(n = ifelse(n > 2, 2, n)) %>% 
  ggplot(aes(n)) +
  geom_histogram(binwidth = 0.05) +
  facet_wrap(~ small, ncol = 1, scales = "free_y") +
  scale_x_continuous(breaks = seq(0, 2, by = 0.1))



tmp10 <-
  tmp %>% 
  #sample_n(1000) %>% 
  mutate(length = as.numeric(st_length(.)) / 1852) %>% 
  st_drop_geometry() %>% 
  left_join(lb %>% select(vid, visir, effort)) 
tmp10 %>% 
  filter(effort < 70000) %>% 
  filter(length < 100) %>% 
  mutate(effort.class = case_when(effort < 21000 ~ "1",
                                  between(effort, 21000, 35000) ~ "2",
                                  TRUE ~ "3")) %>% 
  ggplot(aes(effort, length)) +
  geom_smooth(aes(group = effort.class),
              method = "lm") +
  geom_jitter(alpha = 0.1, size = 0.5, colour = "red") +
  scale_x_log10() +
  scale_y_log10()
tmp10.median <- 
  tmp10 %>% 
  filter(effort < 70000) %>% 
  filter(length < 100) %>% 
  mutate(effort = plyr::round_any(effort, 1000)) %>% 
  group_by(effort) %>% 
  summarise(n = n(),
            median = median(length))
tmp10.median %>% 
  filter(n > 20) %>% 
  ggplot(aes(effort, median)) +
  geom_smooth(method = "lm") +
  geom_point() +
  expand_limits(x = 0, y = 0)
tmp10 %>% 
  filter(effort < 70000) %>% 
  filter(length < 40) %>% 
  ggplot(aes(effort, length)) +
  geom_jitter(alpha = 0.1, size = 0.5, colour = "red") +
  geom_point(data = tmp10.median %>% filter(n > 15), 
             aes(effort, median), colour = "blue", size = 2)


tmp10 %>% 
  filter(effort < 70000) %>% 
  filter(length < 100) %>% 
  ggplot(aes(effort, length)) +
  geom_hex()
tmp10 %>% 
  count(vid, effort) %>% 
  arrange(n)
  
```


```{r}
knitr::opts_chunk$set(eval = FALSE)
```



# INTERPOLATIONS


```{r}
# before interpolation, filter out whacky points
```

```{r}
grid <- 
  expand_grid(vid = unique(lb$vid),
                    time = seq(ymd_hms("2021-01-01 00:00:00"),
                               ymd_hms("2021-12-31 23:59:00"),
                               by = "min")) %>% 
  arrange(vid, time)
```


## HERE THINGS GO WRONG

### VID 2817

```{r}
VID <- 2817
stk.tmp %>% 
  filter(vid == VID) %>% 
  arrange(fishing) %>% 
  st_as_sf(coords = c("lon", "lat"),
           crs = 4326) %>% 
  mapdeck() %>% 
  add_scatterplot(fill_colour = "fishing",
                  radius = 100,
                  palette = "inferno")
stk.tmp %>% 
  filter(vid == VID) %>% 
  ggplot(aes(time, speed)) +
  geom_vline(xintercept = lb %>% filter(vid == VID) %>% pull(t2), colour = "blue") +
  geom_point(size = 0.2, colour = "red")
grd.tmp <- 
  grid %>% 
  filter(vid == VID) %>% 
  left_join(stk.raw %>% filter(vid == VID) %>% distinct(time, .keep_all = TRUE),
            by = c("vid", "time")) %>% 
  arrange(vid, time) %>% 
  group_by(vid) %>% 
  mutate(y = 1:n(),
         lon = approx(y, lon, y, method = "linear", rule = 1, f = 0, ties = mean)$y,
         lat = approx(y, lat, y, method = "linear", rule = 1, f = 0, ties = mean)$y,
         speed = approx(y, speed, y, method = "linear", rule = 1, f = 0, ties = mean)$y) %>% 

grd.tmp %>% 
  #filter(!is.na(lon)) %>% 
  ggplot(aes(time, speed)) +
  geom_vline(xintercept = lb %>% filter(vid == VID) %>% pull(t2),
             colour = "blue") +
  geom_point(size = 0.2, colour = "red")


tmp2 %>% 
  filter(!is.na(lon)) %>% 
  st_as_sf(coords = c("lon", "lat"),
           crs = 4326) %>% 
  mapdeck() %>% 
  add_scatterplot(fill_colour = "speed",
                  radius = 100,
                  palette = "inferno")
```


### VID 2739

```{r}
VID <- 2739
tmp %>% 
  filter(vid == VID) %>% 
  arrange(fishing) %>% 
  st_as_sf(coords = c("lon", "lat"),
           crs = 4326) %>% 
  mapdeck() %>% 
  add_scatterplot(fill_colour = "fishing",
                  radius = 500,
                  palette = "inferno")



tmp2 <- 
  grid %>% 
  filter(vid == VID) %>% 
  left_join(stk.raw %>% filter(vid == VID) %>% distinct(time, .keep_all = TRUE),
            by = c("vid", "time")) %>% 
  arrange(vid, time) %>% 
  group_by(vid) %>% 
  mutate(y = 1:n(),
         lon = approx(y, lon, y, method = "linear", rule = 1, f = 0, ties = mean)$y,
         lat = approx(y, lat, y, method = "linear", rule = 1, f = 0, ties = mean)$y,
         speed = approx(y, speed, y, method = "linear", rule = 1, f = 0, ties = mean)$y) %>% 
  ungroup()
tmp2 %>% 
  filter(!is.na(lon)) %>% 
  ggplot(aes(time, speed)) +
  geom_vline(xintercept = lb %>% filter(vid == VID) %>% pull(t2),
             colour = "blue") +
  geom_point(size = 0.2, colour = "red")


tmp2 %>% 
  filter(!is.na(lon)) %>% 
  st_as_sf(coords = c("lon", "lat"),
           crs = 4326) %>% 
  mapdeck() %>% 
  add_scatterplot(fill_colour = "speed",
                  radius = 100,
                  palette = "inferno")
```



```{r}


d <-
  grid %>% 
  left_join(stk.raw) %>% 
  group_by(vid) %>% 
  mutate(y = 1:n()) %>% 
  rb_interval_id2(lb, vid, time, t1, t2, visir) %>% 
  #filter(vid %in% VIDs[11:20]) %>% 
  mutate(fishing = ifelse(!.id, TRUE, FALSE))
d





d %>% count(vid)
tmp <- 
  d %>% 
  ungroup() %>% 
  filter(vid == 2739) %>% 
  mutate(lon = approx(y, lon, y, method = "linear", rule = 1, f = 0, ties = mean)$y,
         lat = approx(y, lat, y, method = "linear", rule = 1, f = 0, ties = mean)$y,
         speed = approx(y, speed, y, method = "linear", rule = 1, f = 0, ties = mean)$y) %>% 
  rb_interval_id2(lb %>% filter(vid == 2739), vid, time, t1, t2, visir)
tmp %>% 
  filter(!is.na(lon)) %>% 
  st_as_sf(coords = c("lon", "lat"),
           crs = 4326) %>% 
  mapdeck() %>% 
  add_scatterplot(fill_colour = "speed",
                  palette = "inferno")
```

