---
title: "Trips"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE,
  collapse = TRUE,
  error = TRUE,
  comment = "#>"
)
```

## Comparison on trip calculation

```{r}
library(WKSSFGEO)
library(data.table)
library(tidyverse)
```


Jeppe:

```{r}
system.time(d.jp <-
              WKSSFGEO::track01 %>% 
              setDT() %>% 
              unique(by = c("vessel_id", "time_stamp")) %>% # Remove duplicates
              mutate(vessel_id = str_sub(vessel_id, 4) %>% as.integer(),
                     .rid = 1:n()) %>% 
              add_harbours(WKSSFGEO::harbours) %>% 
              define_trips_pol(min_dur = 0.0001, max_dur = 50e6, 
                               split_trips = FALSE, preserve_all = TRUE))
```

Einar:

```{r}
system.time(d.eh <- 
              WKSSFGEO::track01 %>% 
              distinct(vessel_id, time_stamp, .keep_all = TRUE) %>% 
              mutate(vessel_id = str_sub(vessel_id, 4) %>% as.integer(),
                     .rid = 1:n()) %>% 
              sf::st_as_sf(coords = c("lon", "lat"),
                           crs = 4326,
                           remove = FALSE) %>% 
              sf::st_join(WKSSFGEO::harbours %>% mutate(hid = 1:n()) %>% select(hid)) %>% 
              sf::st_drop_geometry() %>% 
              group_by(vessel_id) %>% 
              mutate(inh = !is.na(hid),
                     tid = ramb::rb_trip(inh)) %>% 
              ungroup())


# Pad harbour id, adding harbour id left to the first ping of a trip and harbour id entered
#  to the last ping of the trip.

d.eh %>% 
  group_by(vessel_id) %>% 
  mutate(hid = ramb::rb_pad_harbour(tid, hid)) %>% 
  filter(tid > 0) %>% 
  group_by(vessel_id, tid) %>% 
  slice_tail(n = 1)
```

Number of records:

```{r}
tibble(original = nrow(d),
       jp = nrow(d.jp),
       eh = nrow(d.eh))
```

Any vessel difference?:

```{r}
full_join(d.eh %>% count(vessel_id, name = "n.eh"),
          d.jp %>% count(vessel_id, name = "n.jp"))
```

So vessels 5 and 8 not returned in the jp-workflow.

```{r}
d.eh %>% 
  filter(vessel_id %in% c(5, 8)) %>% 
  count(vessel_id, inh)
```

Reason seemms the jp-flow does not return records if all pings are in harbour. This may be intentional - should possibly be mentioned in the help file that this takes place.

Number of pings per trip:

```{r}
full_join(d.eh %>% mutate(tid = ifelse(tid < 0, NA, tid)) %>% count(vessel_id, tid, name = "n.eh"),
          d.jp %>% separate(trip_id, c("dummy", "tid"), convert = TRUE) %>% count(vessel_id, tid, name = "n.jp")) %>% 
  mutate(diff = n.eh - n.jp) %>% 
  knitr::kable()
```

So difference is in all cases except vessel 7, trip 8 one less ping in the eh-algorithm compared with the jp-algoritm. 

```{r}
tmp <-
  d.eh %>%
  select(.rid, vessel_id, time_stamp, hid.eh = hid, inh, tid.eh = tid) %>%
  full_join(d.jp %>% as_tibble() %>% select(.rid, hid.jp = SI_HARB, tid.jp = trip_id) %>% mutate(in.jp = TRUE)) %>%
  mutate(in.jp = replace_na(in.jp, TRUE)) %>%
  separate(tid.jp, c("dummy", "tid.jp"), convert = TRUE) %>%
  select(-dummy)
```

The reason for the difference of one is that the jp-algorithm includes the first harbour ping at the end of a trip as being part of the trip. Case in point:

```{r}
tmp %>% 
  filter(vessel_id == 1, tid.jp == 1) %>% 
  tail()
```

What about vessel 7, trip 8? 

```{r}
tmp %>% 
  filter(vessel_id == 7, tid.jp == 8) %>% 
  data.table()
```

Seems like the last record for this trip did not terminate in a harbour, as illustrated here (actually it does not even start in a harbour):

```{r}
d.eh %>% 
  filter(vessel_id == 7, .rid > 137770) %>% 
  sf::st_as_sf(coords = c("lon", "lat"),
               crs = 4326) %>% 
  mutate(tid = as.character(tid)) %>% 
  mapview::mapview(zcol = "tid")
```
